---
title: "Match toolmarks with DTW"
author: "Christophe Champod"
date: "04/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load required packages and functions
source("Packages_and_Functions.R")
```

# Load data

The function ImportToolmarksData() below will read all .txt files in a folder and prepare a dataframe with the relevant information.

```{r load_data}
ResToolMarks <- readRDS("ResToolMarks.Rds")
```

A few test using the DTW package. https://dtw.r-forge.r-project.org/
```{r}
a <- Extract_One_profile(ResToolMarks,1)$y
b <- Extract_One_profile(ResToolMarks,35)$y

## Find the best match with the canonical recursion formula
alignment <- dtw(a,b,keep=TRUE)

## Display the warping curve, i.e. the alignment curve
plot(alignment,type="threeway")

## Display distance points (as may as wanted) with offset for viewing
plot(dtw(a,b,k=TRUE), 
          type="two",off=1,match.lty=2,
          match.indices=20)
```

```{r Make_selection}

K1_Z1_A1_ROI1 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 1 & Side == "A1" & ROI == 1)
K1_Z2_A1_ROI1 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 2 & Side == "A1" & ROI == 1)

K1_Z1_A2_ROI1 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 1 & Side == "A2" & ROI == 1)
K1_Z2_A2_ROI1 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 2 & Side == "A2" & ROI == 1)

#Join them up and select y only
Selection1 <- rbind(K1_Z1_A1_ROI1,
                    K1_Z2_A1_ROI1,
                    K1_Z1_A2_ROI1,
                    K1_Z2_A2_ROI1) %>%
  select(y)

#Keep names and design a label to be added to ggdendro
labels_Selection1 <- rbind(K1_Z1_A1_ROI1,
                    K1_Z2_A1_ROI1,
                    K1_Z1_A2_ROI1,
                    K1_Z2_A2_ROI1) %>%
  mutate(Label = paste0("Z",Zone,
                        Side,
                        "ROI", ROI,
                        "/",Cut)) %>%
  select(File,Label)

#If we want the labels on all data
labels_ResToolMarks <- ResToolMarks %>%
  mutate(Label = paste0("Z",Zone,
                        Side,
                        "ROI", ROI,
                        "/",Cut)) %>%
  select(File,Label)
```

We can compute DTW distances for all the entities in Selection1. Note that we can process directly lists with proxy::dist().

## With DTW and dtwclust

```{r DTW_distances}
# To compute DTW pairwise distances
mv_distmat_dtwclust <- proxy::dist(Selection1$y,
                                   method = "dtw_basic")

#To construct a hierarchical clustering
#Distance can be either "dtw" or "dtw_basic" (quicker)
clust.hier <- tsclust(Selection1$y, 
                      type = "hierarchical", 
                      distance = "dtw_basic", 
                      clustering="pam")

#To plot
plot(clust.hier)
```

## DTW with TSclust

Based in the example from https://damien-datasci-blog.netlify.app/post/time-series-clustering-with-dynamic-time-warp/

It gives very similar results compared to tsclust()

```{r TSclust}
#Compute parwise distances
dist_ts <- TSclust::diss(SERIES = Selection1$y, METHOD = "DTWARP")

hc <- stats::hclust(dist_ts, method="complete")
# method can be also "average" or diana (for DIvisive ANAlysis Clustering)
# k for cluster which is 4 in our case 
hclus <- stats::cutree(hc, k = 4) %>% 
  # hclus <- cluster::pam(dist_ts, k = 4)$clustering 
  #has a similar result
  as.data.frame(.) %>%
  dplyr::rename(.,cluster_group = .) %>%
  tibble::rownames_to_column("type_col")

hcdata <- ggdendro::dendro_data(hc)

# Use the following to remove labels from dendogram so not doubling up - but good for checking hcdata$labels$label <- ""

p1 <- hcdata %>%
  ggdendro::ggdendrogram(., rotate=FALSE, leaf_labels=TRUE)

p1
```

It is possible to register a new DTW function adapted to take advantage of specific parameters of the dtw(): step.pattern, open.end, open.begin, etc.

```{r newDTW}
# we specify the new DTW distance as a function
ndtw <- function(x, y, ...) {
  dtw(x, y, ...,
      step.pattern = asymmetric,
      open.end = TRUE,
      open.begin = TRUE,
      distance.only = TRUE)$normalizedDistance
}

#delete name from proxy pr_DB
pr_DB$delete_entry("ndtw")

# Register the distance with proxy
proxy::pr_DB$set_entry(FUN = ndtw, 
                       names = c("nDTW"),
                       loop = TRUE, 
                       type = "metric",
                       distance = TRUE,
                       description = "Normalized, asymmetric DTW")
```

```{r Proxy_Apply_to_selection1}
# Clustering
# It will take some time to compute
cluster.nDTW <- dtwclust::tsclust(Selection1$y,
                      type = "hierarchical",
                      distance = "nDTW")

#Distances
Distance.matrix <- cluster.nDTW@distmat
#Should be the same as
Distance.matrix2 <- proxy::dist(Selection1$y,
                                   method = "nDTW")

#adapt labels - check manually please
cluster.nDTW$labels <- labels_Selection1$Label
#To plot
plot(cluster.nDTW)

#Plot with ggdendro()
ggdendro::ggdendrogram(as.dendrogram(cluster.nDTW) %>%
                         set("labels_cex", 1),
                       rotate=TRUE, labels=TRUE)

#We can get the distances
distmat_dtwclust_nDTW <- proxy::dist(Selection1$y,
                                   method = "nDTW")
```

Test on a selection2 with the three ROIs.

```{r Selection2}
K1_Z1_A1_ROI1 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 1 & Side == "A1" & ROI == 1)
K1_Z1_A1_ROI2 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 1 & Side == "A1" & ROI == 2)
K1_Z1_A1_ROI3 <- ResToolMarks %>%
  filter(Tool == "K1" & Zone == 1 & Side == "A1" & ROI == 3)


#Join them up and select y only
Selection2 <- rbind(K1_Z1_A1_ROI1,
                    K1_Z1_A1_ROI2,
                    K1_Z1_A1_ROI3) %>%
  select(y)

#Keep names and design a label to be added to ggdendro
labels_Selection2 <- rbind(K1_Z1_A1_ROI1,
                    K1_Z1_A1_ROI2,
                    K1_Z1_A1_ROI3) %>%
  mutate(Label = paste0("Z",Zone,
                        Side,
                        "ROI", ROI,
                        "/",Cut)) %>%
  select(File,Label)

# Clustering
# It will take some time to compute
cluster.nDTW.selection2 <- dtwclust::tsclust(Selection2$y,
                      type = "hierarchical",
                      distance = "nDTW")

#adapt labels - check manually please
cluster.nDTW.selection2$labels <- labels_Selection2$Label
#To plot
plot(cluster.nDTW.selection2)

#Plot with ggdendro()
ggdendro::ggdendrogram(as.dendrogram(cluster.nDTW.selection2) %>%
                         set("labels_cex", 1),
                       rotate=TRUE, labels=TRUE)

```

We apply on the all datasets
```{r Proxy_Apply_to_alldata}
# It will take some time to compute
clusterResToolMarks.nDTW <- dtwclust::tsclust(ResToolMarks$y,
                      type = "hierarchical",
                      distance = "nDTW",
                      seed = 1234,
                      trace = TRUE)

clusterResToolMarks.nDTW$labels <- labels_ResToolMarks$Label

#To plot
p2 <- ggdendro::ggdendrogram(as.dendrogram(clusterResToolMarks.nDTW) %>%
                         set("labels_cex", 1),
                       rotate=TRUE, labels=TRUE)
p2

ggsave("dendro1.pdf",p2,width=30,height=100,units="cm")
```

